---
title: "wine_quality_classification"
author: "Yuanyuan Zeng(yz4181)"
date: '2023-02-02'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(gtsummary)
library(ggplot2)
library(corrplot)
library(caret)
```

# Load Dataset
```{r}
#original dataset
wine = read.csv("WineQT.csv") %>% janitor::clean_names() %>% mutate(quality = as.factor(quality))

#copy of dataset
df = wine

#look at data
head(df, 10)
```

# Initial information about the data

## 1. basic information
```{r}
str(df)
```

* The dataset has 13 columns and all columns are numeric. There are no missing value in the dataset.

## 2. descriptive statistics of numeric variables
```{r}
tbl_summary(df)
```

* The variables are not on the same scale.

# Exploratory Data Analysis & Visualization

## Historgram with density curve
```{r}
par(mfrow = c(1,3))
for(i in 1:3) {
  hist(df[,i], freq = FALSE, main = names(df)[i],xlab = names(df)[i])
  lines(density(df[,i]),lwd = 2, col = 4)
}
```

* `fixed_acidity` is relatively normally distributed, while `volatile_acidity` and `citric_acid` have several peaks in the distribution 

```{r}
par(mfrow = c(1,3))
for(i in 4:6) {
  hist(df[,i], freq = FALSE, main = names(df)[i], xlab = names(df)[i])
  lines(density(df[,i]),lwd = 2, col = 4)
}
```

* From the graph above, all distributions are right-skewed.

```{r}
par(mfrow = c(1,3))
for(i in 7:9) {
  hist(df[,i], freq = FALSE, main = names(df)[i], xlab = names(df)[i])
  lines(density(df[,i]),lwd = 2, col = 4)
}
```

** The distribution of `total_sulfur_dioxide` is right-skewed.

Those skewed distributions should be transformed to normal distribution.

## box boxplot
```{r}
colname = names(df[,1:11])

for (i in colname){
  plt = ggplot(data = df, aes_string(y = i, x = "quality", fill = "quality")) +
  geom_boxplot()
  
  print(plt)
  }
```

Some outliers are detected and needed to be removed when cleaning the data. The distributions of residual sugar and chorides are similar for all quality level.

## corrlations between variables
```{r}
corrplot(cor(df[,1:9]), type = "upper", diag = FALSE, method="circle")
```

* correlations of `fixed_acidity` with `citric_acid` and `density` are positive.
* correlations of `free_sulfur_dioxide` with `total_sulfur_dioxide` are positive.

Calculate VIF for variables and remove variables with VIF greater 10.


# Data Cleaning

## duplicate data
```{r}


```





